Predicting the Capture Rate of Pokemon

### 2022-12-11 Final_Draft.Rmd
commit

---
title: "Predicting the Capture Rate of Pokemon"
subtitle: "PSTAT 131 Final Project"
author: "Shivani Kharva"
date: "2022-12-06"
output:
  html_document:
    code_folding: hide
    toc: true
---

@@

This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.

@@

Legendary Pokemon are especially powerful Pokemon that players rarely encounter. From this definition, it seems likely that it would be harder to capture a Pokemon that is legendary than one that is not.

@@

7. Select the best tuned parameters for the model (based on RMSE), adjust the model workflow to include that version of the model, and fit the model with the best tuned parameters to the folds. Make sure to save the final fit.

```{r, eval=FALSE, warning=FALSE, message=FALSE}
# select_best() to select the best tuned parameters for the model (based on RMSE)
# finalize_workflow() to adjust the model workflow to include that version of the model
# fit_resamples() to fit the model with the best tuned parameter to the folds

# LINEAR REGRESSION 
# No need to select the best fit as linear regression had no tuning -> we just fit the one linear regression model we fit
lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
write_rds(lm_fit, file = "data/final_models/lm_final.rds")

# RIDGE REGRESSION
best_penalty <- select_best(ridge_tune, metric = 'rmse')
ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
write_rds(ridge_final_fit, file = "data/final_models/ridge_final.rds")

# LASSO REGRESSION
best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
write_rds(lasso_final_fit, file = "data/final_models/lasso_final.rds")

# POLYNOMIAL REGRESSION
best_poly<- select_best(poly_tuned, metric = "rmse")
poly_final <- finalize_workflow(poly_wf, best_poly)
poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
write_rds(poly_final_fit, file = "data/final_models/poly_final.rds")

# K NEAREST NEIGHBORS
best_knn <- select_best(knn_tune, metric = "rmse")
knn_final <- finalize_workflow(knn_workflow, best_knn)
knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
write_rds(knn_final_fit, file = "data/final_models/knn_final.rds")

# ELASTIC NET
best_elastic <- select_best(elastic_tune, metric = "rmse")
elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
write_rds(elastic_final_fit, file = "data/final_models/elastic_final.rds")

# RANDOM FOREST
best_rf <- select_best(rf_tuned)
rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
write_rds(rf_final_fit, file = "data/final_models/rf_final.rds")

# BOOSTED TREES
best_boosted <- select_best(boosted_tuned)
boosted_final_workflow <- finalize_workflow(boosted_workflow, best_boosted)
boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
write_rds(boosted_final_fit, file = "data/final_models/boosted_final.rds")
```

8. Load in each of the best models and collect the RMSE's of each of the best model fits for comparison with one another.

```{r, warning=FALSE, message=FALSE}
# collect_metrics() to collect the RMSE
# slice() to only collect the RMSE and not R squared

# LINEAR REGRESSION 
lm_final_fit <- read_rds(file = "data/final_models/lm_final.rds")
lm_rmse <- collect_metrics(lm_final_fit) %>% slice(1)

# RIDGE REGRESSION
ridge_final_fit <- read_rds(file = "data/final_models/ridge_final.rds")
ridge_rmse <- collect_metrics(ridge_final_fit) %>% slice(1)

# LASSO REGRESSION
lasso_final_fit <- read_rds(file = "data/final_models/lasso_final.rds")
lasso_rmse <- collect_metrics(lasso_final_fit) %>% slice(1)

# POLYNOMIAL REGRESSION
poly_final_fit <- read_rds(file = "data/final_models/poly_final.rds")
poly_rmse <- collect_metrics(poly_final_fit) %>% slice(1)

# K NEAREST NEIGHBORS
knn_final_fit <- read_rds(file = "data/final_models/knn_final.rds")
knn_rmse <- collect_metrics(knn_final_fit) %>% slice(1)

# ELASTIC NET
elastic_final_fit <- read_rds(file = "data/final_models/elastic_final.rds")
elastic_rmse <- collect_metrics(elastic_final_fit) %>% slice(1)

# RANDOM FOREST
rf_final_fit <- read_rds(file = "data/final_models/rf_final.rds")
rf_rmse <- collect_metrics(rf_final_fit) %>% slice(1)

# BOOSTED TREES
boosted_final_fit <- read_rds(file = "data/final_models/boosted_final.rds")
boosted_rmse <- collect_metrics(boosted_final_fit) %>% slice(1)
```

@@

# Creating a barplot of the RMSE values
ggplot(all_models, aes(x=Model, y=RMSE)) +
  geom_bar(stat = "identity", aes(fill = Model)) +
  scale_fill_manual(values = c("blue1", "red1", "blue2", "red2", "blue3", "red3", "blue4", "red4")) +
  theme(legend.position = "none") 
```

@@

```{r, eval = FALSE, warning = FALSE}
# Fitting to the training data
best_rf_train <- select_best(rf_tuned)
rf_final_workflow_train <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit_train <- fit(rf_final_workflow, data = pokemon_train)

@@

@@

After fitting various models and conducting analysis on each of them, the best model to predict the capture rate of a Pokemon seems to be the random forest model. However, this model still does not do a great job predicting the capture rate (based on the RMSE). This might mean the capture rate of a Pokemon is not very predictable or that the predictors chosen might not have been the best predictors to use.

@@

>>>>

---
title: "Predicting the Capture Rate of Pokemon"
subtitle: "PSTAT 131 Final Project"
author: "Shivani Kharva"
date: "2022-12-11"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

@@

This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon Database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.

@@

Legendary Pokemon are especially powerful Pokemon that players rarely encounter. From this definition, it seems likely that it would be harder to capture a Pokemon that is legendary than one that is not. Let's see if this is true or not.

@@

7. Collect the metrics of the tuned models, arrange in ascending order of mean to see what the lowest RMSE for that tuned model is, and slice to choose only the lowest RMSE. Save the RMSE to a variable for comparison.

```{r, eval=FALSE, warning=FALSE, message=FALSE}
# collect_metrics() to collect the RMSE
# slice() to save only RMSE

# LINEAR REGRESSION 
# Fitting the linear regression to the folds first (since it had no tuning)
lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
lm_rmse <- collect_metrics(lm_fit) %>% 
  slice(1)

# RIDGE REGRESSION
ridge_rmse <- collect_metrics(ridge_tuned) %>% 
  arrange(mean) %>% 
  slice(49)

# LASSO REGRESSION
lasso_rmse <- collect_metrics(lasso_tuned) %>% 
  arrange(mean) %>% 
  slice(35)

# POLYNOMIAL REGRESSION
poly_rmse <- collect_metrics(poly_tuned) %>% 
  arrange(mean) %>% 
  slice(6)

# K NEAREST NEIGHBORS
knn_rmse <- collect_metrics(knn_tuned) %>% 
  arrange(mean) %>% 
  slice(6)

# ELASTIC NET
elastic_rmse <- collect_metrics(elastic_tuned) %>% 
  arrange(mean) %>% 
  slice(73)

# RANDOM FOREST
rf_rmse <- collect_metrics(rf_tuned) %>% 
  arrange(mean) %>% 
  slice(513)

# BOOSTED TREES
boosted_rmse <- collect_metrics(boosted_tuned) %>% 
  arrange(mean) %>% 
  slice(126)
```

@@

# Creating a barplot of the RMSE values
ggplot(all_models, aes(x=Model, y=RMSE)) +
  geom_bar(stat = "identity", aes(fill = Model)) +
  scale_fill_manual(values = c("blue1", "red1", "blue2", "red2", "blue3", "red3", "blue4", "red4")) +
  theme(legend.position = "none") 
```
*Note: the colors in this bar plot do not represent anything and are simply for aesthetic preference.*

@@

```{r, eval = FALSE, warning = FALSE}
# Fitting to the training data
best_rf_train <- select_best(rf_tuned, metric = 'rmse')
rf_final_workflow_train <- finalize_workflow(rf_workflow, best_rf)
rf_final_fit_train <- fit(rf_final_workflow, data = pokemon_train)

@@

We might also be interested in a plot of the predicted values versus the actual values:

```{r}
# Creating plot of predicted values vs. actual values
pokemon_tibble %>% 
  ggplot(aes(x = .pred, y = capture_rate)) +
  geom_point(alpha = 0.4) +
  geom_abline(lty = 2) +
  theme_grey() +
  coord_obs_pred()
```

If every observation was predicted accurately, the dots would form a straight line. From the plot, we can see that very few of the dots actually fall on the line. However, it also appears that the model did not do a terrible job predicting the capture_rate because the points do seem to follow the same general pattern as the line. Also, it is a good sign that the model did not predict any negative capture rate values. So, overall, the random forest model did not do a terrible job predicting the capture rate, but there is definitely room for improvement.

@@

After fitting various models and conducting analysis on each of them, the best model to predict the capture rate of a Pokemon seems to be the random forest model. This is no surprise at all since the random forest tends to work well for most data because it is nonparametric and makes no assumptions about parametric forms or the outcome (more flexible). However, this model still does not do a great job predicting the capture rate (based on the RMSE).

The model that did the worst was K Nearest Neighbors (KNN). This was also not surprising as KNN tends to do worse when there are too many predictors because this means the data has too many dimensions. In a high dimensional data space, the data points are not close enough to each other for KNN to do well in predicting the outcome unless there are enough observations to make up for the high dimensional data space. Therefore, it makes sense that KNN did the worst.

@@

This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon Database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.

The information about Minior was found on [Bulbapedia](https://bulbapedia.bulbagarden.net/wiki/Minior_(Pok%C3%A9mon)), a community Pokemon encyclopedia.

General facts and definitions mentioned in the project were either found or double checked (if from prior knowledge) on the [official Pokemon Database](https://pokemondb.net/). This is also the same database where the data was scraped from.
